---
title: "SOM Plots"
author: "Simon Brewer"
date: "6/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data and libraries

**Make sure that "runSOM.R" has been run first.**
```{r message=FALSE}
library(spdep)
library(kohonen)
library(RColorBrewer)
library(reshape2)
library(ggplot2)

load("largeSOM.RData")
```

## Visualisation

### Change plot
```{r}
plot(poll.som, type="changes", main="Training: SOM 31x30")
```

### Counts plot
```{r}
plot(poll.som, type="counts", main="Counts: SOM 31x30")
```

### Codebook plot
```{r}
plot(poll.som, type="codes", main="Counts: SOM 31x30")
png("codebook.png", width = 4000, height = 4000)
plot(poll.som, type="codes", main="Counts: SOM 31x30", lwd = 3)
dev.off()
```

### Distance plot
```{r}
plot(poll.som, type="dist.neighbours", main="Distances: SOM 31x30")
```

### Age plot

```{r}
## ------------------------------------------------------------------------
myRCBage = function(n, pal="YlGnBu") {
  brewer.pal(n, pal)
}
nnodes = gridX*gridY
nodeAge = rep(NA, nnodes)
grid.dist = unit.distances(poll.som$grid)
for (i in 1:nnodes) {
  nodeID = which(poll.som$unit.classif == i)
  if (length(nodeID) >0) {
    nodeAge[i] = mean(ages[nodeID])
  } else {
    nID = which(grid.dist[i,] < 2)
    nodeID = which(poll.som$unit.classif %in% nID)
    nodeAge[i] = mean(ages[nodeID], na.rm=TRUE)
  }
}

plot(poll.som, type="property", property=nodeAge/1000, main="Mean node age in ka BP",
     palette.name=myRCBage, ncolors=9)
#add.cluster.boundaries(poll.som, poll.som.kmean$cluster)
```

### Clusters

```{r}
plot(som.skat, grid.som$pts, cex.circles=0.035, cex.lab=.7)
```

#### Clusters in SOM space
```{r}
myRCBcls = function(n, pal="Dark2") {
  brewer.pal(n, pal)
}
plot(poll.som, type="property", property=as.numeric(som.clus.ord), main="Cluster map", 
     palette.name=myRCBcls, ncolors=nbclus, heatkeywidth = .5)
add.cluster.boundaries(poll.som, som.clus.ord)
```

#### Clusters in time 
```{r}
plot.df = data.frame(age=nodeAge, cluster=as.factor(som.clus.ord))
# plot.df$cluster.ord = reorder(plot.df$cluster, plot.df$age*-1, FUN=median, na.rm=TRUE)
# plot.df$cluster.ord = som.clus.ord

x = ggplot(plot.df, aes(x=cluster, y=age, fill=cluster)) + 
  geom_boxplot() + scale_fill_brewer(palette="Dark2")
x = x + theme_bw() + scale_x_discrete("Cluster") + scale_y_continuous("Age BP")
x = x + ggtitle("Cluster occurrence by age")
print(x)

x = ggplot(plot.df, aes(x=cluster, y=age, fill=cluster)) + 
  geom_violin() + scale_fill_brewer(palette="Dark2")
x = x + theme_bw() + scale_x_discrete("Cluster") + scale_y_continuous("Age BP")
x = x + ggtitle("Cluster occurrence by age")
print(x)

x = ggplot(plot.df, aes(age, fill=cluster)) + 
  geom_density(alpha=0.5) + scale_fill_brewer(palette="Dark2")
x = x + theme_bw() + scale_x_reverse()
x = x + ggtitle("Cluster occurrence by age")
print(x)
```

#### Abundances by cluster
```{r}
plot.df = cbind(melt(poll.som$codes[[1]], varnames = c("Grid","Taxa"),
                     value.name = "Score"), Cluster=as.factor(rep(som.clus.ord, 22)))

x = ggplot(plot.df, aes(x=Taxa, y=Score, fill=Cluster)) + geom_boxplot() + facet_wrap(~Cluster)
x = x + ggtitle("All abundance values") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(x)
```

### Taxon maps

How do variables map to the SOM? 
```{r}
myRCBpol = function(n, pal="Greens") {
  brewer.pal(n, pal)
}
taxNames = names(poll)
for (i in 1:length(taxNames)) {
  plot(poll.som, type="property", property=poll.som$codes[[1]][,taxNames[i]], main=taxNames[i], 
       palette.name=myRCBpol, ncolors=9)
  # add.cluster.boundaries(poll.som, som.clus.ord)
}
```

### Site trajectories
REMOVED FOR NOW - NEED TO CHECK CODE

```{r}
allsites = unique(sitesn)
length(allsites)
for (i in 1:length(allsites)) {
  siteID = which(sitesn==allsites[i])
  # plot(poll.som, type="dist.neighbours", keepMargins=TRUE, main=allsites[i])
  plot(poll.som, type="property", property=poll.som$codes[[1]][,"Tsuga"],
       main=allsites[i], keepMargins=TRUE,
       palette.name=myRCBpol, ncolors=9)
  # add.cluster.boundaries(poll.som, som.clus)
  site.crd <- poll.som$grid$pts[poll.som$unit.classif,][siteID,]
  lines(site.crd, lwd=2)
  points(jitter(site.crd))
}
```

### Site trajectories by cluster
REMOVED FOR NOW - NEED TO CHECK CODE

```{r}
allsites = unique(sitesn)
length(allsites)
for (i in 1:length(allsites)) {
  siteID = which(sitesn==allsites[i])
  # plot(poll.som, type="dist.neighbours", keepMargins=TRUE, main=allsites[i])
  plot(poll.som, type="property", property=as.numeric(som.clus.ord),
       main=allsites[i], keepMargins=TRUE,
       palette.name=myRCBcls, ncolors=nbclus, heatkeywidth = .5)
  add.cluster.boundaries(poll.som, som.clus)
  site.crd <- poll.som$grid$pts[poll.som$unit.classif,][siteID,]
  lines(site.crd, lwd=2)
  points(jitter(site.crd))
}
```
